---
title: "HW 1 1st Problem"
author: "Lintao Zhang"
date: "2024-09-11"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Working directory set: 
setwd("D:/3201/Week 2/HW 1")


# read in data from file (do not forget to set working directory)

newspaper.df = read.csv(file="pulitzer-circulation-data.csv",
header=T, stringsAsFactor=F)

```
## DATA

```{r}
head(newspaper.df) ##input all the datasets
```


## Change the data here to be numbers by removing any commas and converting columns 
```{r}
newspaper.df$Daily.Circulation..2004 <- as.numeric(gsub(',', '', newspaper.df$Daily.Circulation..2004))
newspaper.df$Daily.Circulation..2013 <- as.numeric(gsub(',', '', newspaper.df$Daily.Circulation..2013))
newspaper.df$Change.in.Daily.Circulation..2004.2013 <- as.numeric(gsub("%", "", newspaper.df$Change.in.Daily.Circulation..2004.2013))

```



## Part A

## The summary statistics for daily circulation in 2004 AND 2013 of the newspapers
```{r}

#Printing results for 2004
#Using the "cat" here("concatenate and print")to make the output looks more good and concise!:)

#to print summary statistics for given years and datas

print_summary = function(year, data) { 
  # Print header

  cat(" \n")
  cat("----------------------------------------\n")

  cat("Summary Statistics for", year, "Daily Circulation\n")
  
  cat("---------------------------------------\n")
  
  #print key
  cat("2 measures of center:" , "\n")
  cat(" \n")
  cat("Mean:", mean(data, na.rm = TRUE), "\n")
  cat("Median:", median(data, na.rm = TRUE), "\n")
  cat(" \n")
  cat("---------------------\n")
  ##the at 2 measures of center
  
  cat(" \n")
  cat("3 measure of dipersion: " , "\n")
  cat(" \n")
  cat("Standard Deviation:", sd(data, na.rm = TRUE), "\n")
  cat("IQR:", IQR(data, na.rm = TRUE), "\n")
  cat("Range:", range(data, na.rm = TRUE), "\n")
  ##the at 3 measures of dispersion
  
  cat("----------------------------------------\n")
  cat(" \n")
  cat(" \n")

}

#print summary statistics for 2004 and 2013 here
print_summary(2004, newspaper.df$Daily.Circulation..2004)

print_summary(2013, newspaper.df$Daily.Circulation..2013)
```
## Based on the result here, I see that there in 2013, the daily circulation of newspaper has decrease since its mean and median both decreased compared to 2004. But I see the standard deviation increased from 2004 to 2013, meaning that there is more variability in newspaper circulation in 2013. I think this may indicate that while some newspapers have declines, others remained more stable or even increased. AND the range widened dramatically from 215,735 to 2,192,098 in 2004 to 0 to 2,378,827 in 2013. I guess the lower bound being 0 in 2013  means that some newspapers might already shut down during it time range.







## Part B

## Fractions of the newspaper had more than 3Pulitzer Prize finalists and winners( 1990-2003, 2004-2014.)
```{r}
#subset for 1990-2003
subset_1990_2003 = newspaper.df[newspaper.df$Pulitzer.Prize.Winners.and.Finalists..1990.2003 > 3, ]

#subset for 2004-2014
subset_2004_2014 = newspaper.df[newspaper.df$Pulitzer.Prize.Winners.and.Finalists..2004.2014 > 3, ]


#fraction for 1990-2003
fraction_1990_2003 = nrow(subset_1990_2003) / nrow(newspaper.df)


#fraction for 2004-2014
fraction_2004_2014 = nrow(subset_2004_2014) / nrow(newspaper.df)

#results
cat("Fraction for 1990-2003 is:", fraction_1990_2003, "\n")
cat("Fraction for 2004-2014 is:", fraction_2004_2014, "\n")
```




## Part C
## Creat histograms for two groups
```{r}
## Base R


## Group 1: All Newspapers
hist(newspaper.df$Change.in.Daily.Circulation..2004.2013, 
     main = "Group 1: All Newspapers",  
     xlab = "Change in Daily Circulation", 
     col = "red", breaks = 30)

## Group 2: Newspapers with More Than 3 Pulitzer Finalists or Winners (2004-2014)
hist(newspaper.df$Change.in.Daily.Circulation..2004.2013
     [newspaper.df$Pulitzer.Prize.Winners.and.Finalists..2004.2014 > 3], 
     main = "Group 2: Newspapers with > 3 Pulitzer Prizes", #5
     xlab = "Change in Daily Circulation", 
     col = "purple", breaks = 30)

##newspaper.df ## datasets
```


## Part D
## Creat the plot
```{r}

##make some plot of Pulitzer Finalists/Winners vs. Change in Circulation (2004-2014)

plot(newspaper.df$Pulitzer.Prize.Winners.and.Finalists..2004.2014, newspaper.df$Change.in.Daily.Circulation..2004.2013, 
     main = "Finalists/Winners vs. Change in Circulation (2004-2014)", 
     xlab = "Number of Pulitzer Finalists/Winners (2004-2014)", ylab = "Change in Circulation 2004-2014", 
     col = "blue", 
     pch = 21 ) ##blue circle with empty middle


##fit a simple linear model here
model = lm(Change.in.Daily.Circulation..2004.2013 ~ Pulitzer.Prize.Winners.and.Finalists..2004.2014, 
            data = newspaper.df) ##data

#line
abline(model, col = "red")



```

## The scatterplot shows a weak positive correlation between the number of Pulitzer finalists/winners and changes in circulation from 2004 to 2014, with a slight upward trend in circulation for newspapers with more Pulitzer recognition. 
